\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{hyperref}

\title{Handoff Summary: The Diffeology Archives Project (Window 4)}
\author{Patrick Iglesias-Zemmour \& Gemini}
\date{December 25, 2025}

\begin{document}

\maketitle

\section{Session Overview}
This session (Christmas 2025) was a major expansion of the archive. We moved from establishing workflows to archiving foundational historical documents and cutting-edge preprints. We also refined the theoretical narrative of the repository (Erlangen Program vs. Boman Paradox).

\section{Work Completed}

\subsection{Infrastructure}
\begin{itemize}
    \item \textbf{Project Log:} Established \texttt{/Project-Log/} for maintaining these summaries and the archival methodology.
    \item \textbf{Root README:} Updated with "Technical Standards" (MacTeX 2025 compatibility) and Credits.
    \item \textbf{Macro Library:} Updated \texttt{diffeology-piz.sty} to \textbf{v2.13}.
\end{itemize}

\subsection{Archived Papers}
\begin{enumerate}
    \item \textbf{1986-Connexions-et-Diffeologie:}
    \begin{itemize}
        \item \emph{Status:} Unpublished Preprint (CPT-86/P.1931).
        \item \emph{Action:} Full transcription from PDF scan to \LaTeX. Recreated diagrams with \texttt{tikz-cd}.
        \item \emph{Note:} Documented the "Sorites Incident" (rejection by Dieudonné) in the README to explain the historical gap.
    \end{itemize}
    \item \textbf{2010-Orbifolds-as-Diffeologies:}
    \begin{itemize}
        \item \emph{Action:} Transmuted to \texttt{amsart}. Recreated legacy \texttt{eepic} figures in \texttt{TikZ}.
    \end{itemize}
    \item \textbf{2019-Differential-Forms-On-Manifolds-With-Boundary-and-Corners:}
    \begin{itemize}
        \item \emph{Action:} Transmuted to \texttt{amsart}. Cleaned preamble.
    \end{itemize}
    \item \textbf{2021-An-Introduction-To-Diffeology:}
    \begin{itemize}
        \item \emph{Action:} Transmuted. Added "Significance" section regarding the shift in geometric perspective.
    \end{itemize}
    \item \textbf{2025-Bosonic-and-Fermionic-Singularities-in-Diffeology:}
    \begin{itemize}
        \item \emph{Status:} Draft Preprint.
        \item \emph{Action:} Archived source. Added "Singular Capacity" concept to README.
    \end{itemize}
\end{enumerate}

\subsection{Theoretical Linking}
\begin{itemize}
    \item Updated \textbf{2025-The-Boman-Paradox} to explicitly state it invalidates the strict Erlangen view.
    \item Cross-referenced this paradox in the \textbf{2021 Introduction} to show the evolution of the theory.
\end{itemize}

\section{Next Steps}
\begin{itemize}
    \item \textbf{Thesis:} Archive the 1985 State Doctorate Thesis \textbf{"Fibrations difféologiques et Homotopie"}.
\end{itemize}

\end{document}
