

  \ \vfill 

  \begin{center}
   \textbf{Thanks}
  \end{center}
  
  \bigskip
  
  \begin{text}
  Thanks to Yael Karshon and Fran\c{c}ois Ziegler for their 
  encouragements to write this textbook on diffeology. 
  I am especially grateful to Yael Karshon for our many discussions on orbifolds
  and smooth structures,
  and particularly for our discussions on \guillemots{symplectic diffeology} which pushed me to think 
  seriously about this subject.
    
  I am very grateful to The Hebrew University of Jerusalem 
  for its hospitality, specials thanks to  Emmanuel Farjoun 
  for his many invitations to work there. 
  I started, and mainly carried out, this project in Jerusalem. 
  Without my various stays in Israel I would never have finished it.
  Givat Ram campus is a wonderful place where I enjoyed the studious and   
  friendly atmosphere, and certainly its peaceful and intense scientific environment. 
  Thanks also to David Blanc from the University of Haifa, 
  to Misha Polyak and Yoav Moriah from the Technion, to Misha Katz from 
  Bar-Ilan University, to Leonid Polterovich and Paul Biran, from Tel-Aviv University, 
  for their  
  care and invitations to share my ideas with them.
  Thanks also to  Yonatan Barlev,  Erez Nesrim and Daniel 
  Shenfeld, from the Einstein Institute, for their participation in 
  the 2006 seminar on Diffeology. Working together
  has been very pleasant and  their remarks 
  have been helpful. 
  
  Thanks to Paul Donato, my old accomplice in diffeology, 
  who checked some of my claims when I needed a second look. 
  Thanks also to the young guard: Jordan Watt, Enxin Wu and Guillaume Tahar, their 
  enthusiasm is refreshing and their contribution somehow significant.
  Thanks also to the referee of the manuscript for his support and his 
  clever suggestions
  which made me improve the content. 
  
  Thanks, obviously, to the CNRS that pays my salary, and gives me the freedom  
  necessary to achieve this project. Thanks particularly
  to J\'er\^ome Los for his support when I needed it. 
  
  Thanks also to Liliane P\'erichaud who spent hours to help fixing my
  misprints, and to David Trotman for trying to make my English look English.
\end{text}

\begin{center}

    \medskip 
    Jerusalem, September 2005 --- Aix-en-Provence, August 2012
    \bigskip

    Laboratoire d'Analyse, Topologie et Probabilit\'es du CNRS \\
    39, rue F. Joliot-Curie \\
    13453 Marseille, cedex 13 \\
    France \\
    \smallskip \& \\ \smallskip 
    The Hebrew University of Jerusalem \\
    Einstein Institute \\ Campus Givat Ram \\
    91904 Jerusalem\\ Israel \\
    \medskip
    {\tt http://math.huji.ac.il/\hspace{-0pt}\raisebox{-3pt}{\~{}}piz/}\\
    {\tt piz@math.huji.ac.il} 

\end{center}


\vfill
